---
title: "Homework 1 - Statistics 20"
output: pdf_document
---
\begin{center}
\textit{Due April 8, 2016 before 11:59 PM via upload to CCLE}
\end{center}

\section{Installing R \& RStudio}

\textbf{Problem 1.} Please install the most recent versions of R (3.2.1) & RStudio on your computer. As proof of completion, issue this commnd if you use MacOS:

```{r,comment=NA}
system("who", intern = TRUE)
```

but if you use Windows:

```{r,comment=NA}
system("whoami", intern = TRUE)
```

\section{Swirl Package}

\textbf{Problem 2.} In this problem, we'll be using the swirl software package for R in order to practice some key concepts. The swirl package turns the R console into an interactive learning environment.

- Install swirl using:  **install.packages("swirl")**
- Load Swirl using: **library(swirl)**
- Install the R Progroamming course using: **install_from_swirl("R Programming")**
- Start swirl and complete the lessons using: **swirl()**
- Do the following lessons:

A. Do  *1: R Programming Basic Building Blocks* lesson. Explain the *recycling* rule with an example.

When given two vectors of the same length, R simply performs the specified arithmetic operation (`+`, `-`,
 `*`, etc.) element-by-element. If the vectors are of different lengths, R 'recycles' the shorter vector until it is the same length as the longer vector.
e.g.
```{r,comment=NA}
z<-c(1.1,9,3.14)
z*2+100
```
When we do z * 2 + 100, z is a vector of length 3, but technically 2 and 100 are each vectors of length 1. So R is 'recycling' the 2 to make a vector of 2s and the 100 to make a vector of 100s.

B. Do  *3: Sequences of Numbers* lesson. What command was suggested for generating a vector to contain 10 zeros, then 10 ones, then 10 twos?
```{r,comment=NA}
rep(c(0,1,2),each=10)
```

C. Do *4: Vectors* lesson. What was the output for paste(my_name,collapse=" ")?
```{r,comment=NA}
my_char<-c("My","name","is")
paste(my_char, collapse = " ")
```

D. Do *7: Matrices and Data Frames*. What was the output for colnames(my_data)?
```{r,comment=NA}
my_vector<-1:20
dim(my_vector) <- c(4, 5)
my_matrix<-my_vector
patients<-c("Bill", "Gina", "Kelly", "Sean")
cbind(patients,my_matrix)
my_data <- data.frame(patients, my_matrix)
cnames<-c("patient", "age", "weight", "bp", "rating", "test")
colnames(my_data)<-cnames
colnames(my_data)
```


\subsubsection{Character (String) and Logical Vectors}

\textbf{Problem 3.} Suppose a student has written the following:

```{r,comment=NA}
a <- c("Jane", "Dave", "Ann")
b <- c(TRUE, FALSE, FALSE)
b+5
```

```{r, error=TRUE,comment=NA}
d <- matrix(c(a,b), nrow=3)
d[,2]+5
```

Why does d[,2]+5 result in an error but b+3 did not? This answer only requires an explanation, no code.

When we calculate b+3, R automatically converts the logical value into a numeric, so we are acturally calculating c(1,0,0)+5. However, when we combine a and b into a matrix, R automatically converts the logical values into characters, so we are acturally calculating c("TRUE", "FALSE", "FALSE")+5, which is not allowed.
\subsection{Matrices}

\textbf{Problem 4.} For a matrix

```{r,comment=NA}
X <- matrix(56:100, ncol=9)
X
```

A. Show how to display the values in the 2nd column only.
```{r,comment=NA}
X[,2]
```

B. Show how to display the values in 3rd and 5th rows only.
```{r,comment=NA}
X[3,]
X[5,]
```

\subsection{Data Frames}

\textbf{Problem 5.} Data frames share some of the properties of matrices and of lists. Technically, they are considered lists (generic vectors) in the R language. They are used as the fundamental data structure by most of R’s estimation functions.

A. Demonstrate that you are able to use the function data() and that you know what a data frame is by finding a built-in dataset that is NOT USArrests. Please identify the dataset. For example:

```{r,comment=NA}
data() #show all the built-in datatsets
data(BOD)
```

Then do the following:

B. Issue the class() and mode() functions for your data frame so that we can see you truly selected an appropriate dataset.

```{r,comment=NA}
class(BOD)
```

```{r,comment=NA}
mode(BOD)
```

C. How many observations and how many variables does your data frame have? Use any R function to reveal this information.
```{r,comment=NA}
nrow(BOD) #number of observations
ncol(BOD) #number of variables
```

D. Apply the summary function to a column of your choice.
```{r,comment=NA}
summary(BOD[,2])
```

E. Apply the head function to a column of your choice.
```{r,comment=NA}
head(BOD[,2])
```

